\subsection{Chosen Job Advertisement}

\textbf{Polytec Personnel Ltd}

We are looking for a Software Engineer to join a company just outside of Cambridge to work on a wide variety of interesting and challenging projects. You will be involved in all aspects of the software life cycle, from design, to implementation, to testing and maintenance.

Requirements

\begin{itemize}
	\item Degree in Computer Science or related field (2:1 or above)
	\item Experience with one or more development languages (C, C-Sharp, C++, ...)
	\item Proven numerical and analytical skills
	\item Keen eye for detail and problem-solving ability
\end{itemize}

The following would be beneficial: 

\begin{itemize}
	\item Development of embedded real-time systems in C/C++
	\item Data acquisition and analysis
	\item Control systems
	\item CAN, Ethernet, FlexRay
	\item Embedded Linux development in C++
\end{itemize}

Please get in touch for further details or to apply.

\subsection{The Cover Letter}

\textit{This cover letter contains made up information.}

\newpage
\textbf{Software Engineer (Embedded)}

Otto Martinwall

+46 70 598 59 05

otto@martinwall.se

April 11, 2022
\bigskip

Dear Hiring Manager,

I am excited to be able to apply for the open Software Engineer position that you now have. I have a masters degree in Technical Physics and I have therefore been training my problem solving, numerical and analytical skills to a very high degree. I have extensive experience with the languages C\#, JavaScript and Python, which should align well with your requirements. In addition, I enjoy working with networking, servers and small embedded systems as a hobby at home, so I am familiar with Linux/GNU, CAN, Ethernet and FlexRay.

In my previous work at Logiwaste in Sweden, I worked with embedded control systems for garbage suction. Good structure and scalability was and will be of top priority. I believe that my knowledge and experience from that work will transfer well to the projects at your company.

In my free time, I passionately enjoy working with mathematics and physics. I also frequently go running, preferably in the forest, and I value my relationships very highly so I spend much time with my family and friends.

Thank you for your time and consideration. I am looking forward to learning more about the position that you have open at your company and about the work that you are doing.

Yours faithfully



































