\subsection{Literary Elements}

\subsubsection{Characterization}

The protagonist struggles with his psyche. He often feels lonely and angry, which is highly understandable.

\textit{"I'm an animal! he exulted. I'm a dumb, stupid animal and I'm going to drink! He emptied the glass, then flung it across the room. It bounced off the bookcase and rolled across the rug. Oh, so you won't break, won't you! he rasped inside his head, leaping across the rug to grind the glass into splinters under his heavy shoes."}

He also experiences need for human touch, after being alone for so long. He seems to be a terrible person though, which is seen in this quote when talking about the thought-to-be human woman that he brought home to him by force:

\textit{"The most unusual feature ofthe entire affair, he thought, was that he felt no phsysical desire for her. If she had come two years before, maybe even later, he might have violated her."} p.74

or in this other quote where he is beside a currently disabled/sleeping vampire woman during the day:

\textit{"He took the woman from her bed, pretending not to notice the question posed in his mind: why do you always experiment on women? He didn't care to admit that the inference had any validity. She just happened to be the first one he'd come across, that was all. What about the man in the living room, though? For God's sake! he flared back. I'm not going to rape the woman!"} p.32

and this a bit later on the same page about the same woman:

\textit{"Makes a god excuse, doesn't it, Neville? Oh, shut up. But he wouldn't let himself pass the afternoon near her. After binding her to a chair, he secluded himself in the garage and puttered around with the car. She was wearing a torn black dress and too much was visible as she breathed. Out of sight, out of mind.... It was a lie, he knew, but he wouldn't admit it.}

By the end of the book, he finds peace in his death. He was to be executed by a new society of living vampires, but he got a fatal pill by a woman vampire out of kindness, the same woman vampire which he kidnapped before, which he then took before the execution.

\subsubsection{Setting}

The story takes place in Los Angeles, 1976. The book is written to take place in the future, since it was written during the 1950s.

\subsubsection{Plot}

\begin{itemize}
	\item \textbf{\textit{Beginning/Exposition}:} Robert Neville is stuck in Los Angeles, 1976, probably as the lone survivor of the apocalypse where everyone turned into vampires.
	\item \textbf{\textit{Major Conflict}:} The apocalypse.
	\item \textbf{\textit{Minor Conflicts}:} Not commiting suicide, understand and find a cure to the bacteria, find others that are still living.
	\item \textbf{\textit{Rising Action}:} Slow descent into madness and loneliness. Thoughts about suicide. Finding a still living dog and watching it die. Eventually finding another human which seems to be a woman at first.
	\item \textbf{\textit{Climax}:} After the woman knocks Robert out, she lefts a note that reveals that she is a vampire, and that the still living vampires have created a new society. 
	\item \textbf{\textit{Resolution}:} Robert Neville is captured by this new society and executed.
\end{itemize}

\subsubsection{Themes}

\begin{itemize}
	\item \textbf{\textit{Loneliness}:} A constant topic in this book is the protagonist's depression. 
	\item \textbf{\textit{Survival}:} The protagonist is in a world where survival is not easy, and much of earlier parts of the book is spent with describing the things he does to keep living and to keep his house functioning.
	\item \textbf{\textit{Science}:} The protagonist can't accept that the vampire is something supernatural, so he begins  studying biology and bacteria to find explanations for the vampires' behavior, and to maybe be able to find a cure.
\end{itemize}

\subsubsection{Style}

The writing style feels very direct and is easily read. You do not have to decipher much.

\subsubsection{Tone}

The tone is quite dramatic and serious.

\subsubsection{Point of View}

This novel uses a third person limited omniscient narrator.

